A formal approach but practical approach in train regulation

### 2025-06-26 paper/main.tex
composition, structure

% Je suis assez d'avis de distinguer la direction des blocks et des trains
\newcommand{\forward}{\ensuremath{\mathtt{f}}}
\newcommand{\backward}{\ensuremath{\mathtt{b}}\xspace}
\newcommand{\sucblock}{{\mathtt{suc}}}

@@

\begin{document}
%
\title{A formal approach but practical approach in train regulation}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Lucas Villaume \and Guillaume Bonfante \and Martin Vassor}
%
\authorrunning{L. Villaume et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Université de Lorraine, CNRS, Inria, LORIA, F-54000 Nancy, France}
%
\maketitle
%
\begin{abstract}
	% Contexte
	Trains networks are critical system which failure may lead to numerous fatalities, in case of crashes, or important loss, if trains are not properly routed. 
	% Problème
	In this paper, our goal is to formally verify the correctness of train routes. More precisely, we verify the absence of deadlocks, the termination, and the absence of crashes.
	% Approche
	To address this question, we model train systems on paper and in TLA+. Our model includes trains, as well as switches, detection sensors and traffic lights. Both switches and traffic lights are controlled by a regulator upon train events. We provide trains and the regulator with orders, characterising the schedule, which is then checked using TLA+'s model checker.
	% Resultats
	Our contributions are threefold: first, we provide formal model of railroad networks and schedules; second, we implement this model as a TLA+ specification and we report tests on checking schedules on a non-trivial network, composed of 3 trains, 8 blocks and 5 switches; finally we show how to sequentially compose schedules, which we check against our TLA+ specification.

\keywords{Train routing \and Model checking \and Formal methods \and TLA+}
\end{abstract}

@@

\paragraph{Motivation.}

\paragraph{Related works.}

\paragraph{N-scale model.}
\begin{wrapfigure}[10]{r}{0.28\textwidth}
	\begin{minipage}{0.28\textwidth}
		\centering
		\vspace{-8mm}
		\includegraphics[height=30mm]{maquette.png}
		\caption{The N-scale model.}
	\end{minipage}
\end{wrapfigure}
We tested the work presented in this paper on a real-world N-scale train model. Our circuit has 8 blocks, 5 switches. Each block is bidirectional with a signal at each end. Blocks size goes from 25cm to 70cm. We operate three trains concurrently. Due to the size of the model, we reduced our signals  to "red" and "green" status, the trains being too close one from the other. Furthermore, in case of a red signal, we make the hypothesis that the train are able to stop before reaching the end of the block.

\paragraph{Outline.}
In Section~\ref{sec:model_configurations}, we introduce the main building blocks of our model: network infrastructure, trains and the regulator. In Section~\ref{sec:model_semantics}, we present the dynamic aspects of our system: how trains can move and how the regulator schedules them. We conclude this section by stating what we expect of correct networks. Then, in Section~\ref{sec:tla-formalisation}, we present a TLA+ implementation of our model, which we use to check network correctness: model checking a schedule \emph{certifies} its correctness. Our last contribution, in Section~\ref{sec:composition}, explains how to build large schedules by composing smaller ones. While we don't prove the correctness \emph{by construction} of composed schedules, we use the TLA+ implementation to certify, \emph{a posteriori}, the correctness of composed schedules.

@@

\paragraph{Experiments}  With TLA+, we checked that any safe step is actually (not surprisingly!) reliable.  On our network, one finds 58 368 safe steps, all being reliable. Checking all the path took XXX seconds on YYY. \todo{XXX= nombre de seconde, YYY = décrire la machine (processeur, RAM, etc)}

@@

So, we suppose we are given safe step states $S_1, \ldots, S_k$ corresponding to $X_1, \ldots, X_k$. To compute their composition $S_1 \cdot S_2 \cdots S_k$, we proceed by induction on $k$. For $k = 1$, we return $S_1$. For $k>0$, let $S = S_1 \cdots S_{k-1}$. Let $S = \langle \guardG, \bufTrain, \bufSig, \langle \Gamma, \regulator, \switches, \signals\rangle \rangle$. We set $S \cdot S_k = \langle \guardG, \bufTrain, \bufSig, \langle \Gamma', \regulator', \switches, \signals\rangle \rangle$ as follows. First,  for any train $\tid{i} \in \trains$, let $P_0 = S.\Gamma[\tid{i}].P$ and $P_k = S_k.\Gamma[\tid{i}].P$, then set $\Gamma'[\tid{i}] = \Gamma[\tid{i}][ P \leftarrow P_0 \cdot P_k]$.

Second, to define $E'$, we start $E' = E$ and we enrich it as follows.  For each

>>>>

% Je suis assez d'avis de distinguer la direction des blocks et des trains
\newcommand{\forward}{\ensuremath{\mathtt{f}}}
\newcommand{\backward}{\ensuremath{\mathtt{b}}\xspace}
\newcommand{\sucblock}{{\mathtt{suc}}}
\newcommand{\throughS}{{\mathtt{through}}}

@@

\begin{document}
%
\title{A formal approach but practical approach in train regulation}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Lucas Villaume \and Guillaume Bonfante \and Martin Vassor  }
%
\authorrunning{L. Villaume et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Université de Lorraine -- CNRS -- INRIA -- LORIA}
%
\maketitle
%
\begin{abstract}
	% Contexte
	Trains networks are critical system which failure may lead to numerous fatalities, in case of crashes, or important loss, if trains are not properly routed. 
	% Problème
	In this paper, our goal is to formally verify the correctness of train routes. More precisely, we verify the absence of deadlocks, the termination, and the absence of crashes.
	% Approche
	To address this question, we model train systems in TLA+. Our model includes trains, as well as switches, detection sensors and traffic lights. Both switches and traffic lights are controlled by a regulator upon train events. We provide trains and the regulator lists of event, which characterise the train scheduling, which is then checked using TLA+'s model checker.
	% Resultats
	Our contributions are threefold: first, we provide a TLA+ specification which can be adapted to various train networks and schedules; second, we report tests we ran on the model on a non-trivial network, composed of 3 trains, 8 blocks and 5 switches; finally we show how to sequentially compose two schedules and to verify the result.

\keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}

@@

\paragraph{Outline.}
In Section~\ref{sec:model_configurations}, we introduce the main building blocks of our model: network infrastructure, trains and the regulator. In Section~\ref{sec:model_semantics}, we present the dynamic aspects of our system: how trains can move and how the regulator schedules them. We conclude this section by stating what we expect of correct networks. Then, in Section~\ref{sec:tla-formalisation}, we build a TLA+ implementation of our model and its dynamics, which we use to check network correctness. Thus, model checking a schedule \emph{certifies} its correctness. Our last contribution, in Section~\ref{sec:composition}, explains how to build large schedules by composing smaller ones. While we don't prove the correctness \emph{by construction} of composed schedules, we use the TLA+ implementation to certify, \emph{a posteriori}, the correctness of composed schedules.

@@

\paragraph{Experiments}  With TLA+, we checked that any safe step is actually (not surprisingly!) reliable.  On our network, one finds 58 368 safe steps, all being reliable. Checking all the paths took XXX seconds on YYY. \todo{XXX= nombre de seconde, YYY = décrire la machine (processeur, RAM, etc)}

@@

So, we suppose we are given safe step states $S_1, \ldots, S_k$ corresponding to $X_1, \ldots, X_k$. To compute their composition $S_1 \cdot S_2 \cdots S_k$, we proceed by induction on $k$. For $k = 1$, we return $S_1$. For $k>0$, let $S = S_1 \cdots S_{k-1}  = \langle \guardG, \bufTrain, \bufSig, \langle \Gamma, \langle E, W, T\rangle, \switches, \signals\rangle \rangle$. We set $S \cdot S_k = \langle \guardG, \bufTrain, \bufSig, \langle \Gamma', \langle E', W, T\rangle, \switches', \signals\rangle \rangle$ as follows.

First,  for any train $\tid{i} \in \trains$, let $P_0 = S.\Gamma[\tid{i}].P$ and $P_k = S_k.\Gamma[\tid{i}].P$, then set $\Gamma'[\tid{i}] = \Gamma[\tid{i}][ P \leftarrow P_0 \cdot P_k]$.

Second, we proceed by updates of $E$ and $\sigma$. At first, we set $E' = E$ and $\sigma' = \sigma$. Now, suppose that $\sigma_k$ is the network configuration coming with $X_k$. Let $\sigma_0 = \{ \langle \sid{i}, d_0, d_k \rangle \mid \sigma(\sid{i}) = d_0 \neq d_1 = \sigma_k(\sid{i})\}$. For each such triple $\langle \sid{i}, d_0, d_k\rangle$, either there is a step $j < k$, a train $\tid{m}$ such that $\sid{i} \in \throughS(x_n, d, s, x_{n+1})$ with $\vec{x}$ being the route of train $\tid{m}$ at step $X_j$, in which case we keep the quadruple with the highest $j$. If there is no such quadruple, set $\sigma' := \sigma'[ \sid \leftarrow d_k]$.  Then, set $E' := E'[ ....]$.
